Technical Report

### 2017-09-28 fully-abstract-comp/technical-report.tex
changes to the semantics related to our wednesday discussions.

\newcommand{\scbnf}{\shareddom{sc}}
\newcommand{\cbnf}{\shareddom{c}}
\newcommand{\permbnf}{\shareddom{perm}}
\newcommand{\addrbnf}{\shareddom{a}}
\newcommand{\basebnf}{\shareddom{base}}
\newcommand{\aendbnf}{\shareddom{end}}
%\newcommand{\glbnf}{\shareddom{gl}}
\newcommand{\linbnf}{\shareddom{l}}
\newcommand{\sealbasebnf}{\sigma_\shareddom{base}}
\newcommand{\sealendbnf}{\sigma_\shareddom{end}}

@@

%%% Constants
\newcommand{\constant}[1]{\mathrm{#1}}
\newcommand{\calllen}{\constant{call\_len}}

@@

% \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
% \lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
% \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf,\sigma}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma) \mid \retptrc(\sigma)}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{06-09-2017: Should seals also have the option to be linear?}
\lau{13-09-2017: For now I have added linearity to the seals.}
\dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
\lau{28-09-2017: Why would you like this? If we want to add this, we will have to be careful, because an old code-return-pointer may fit perfectly well with our new data-return-pointer, so in this case it should not fail. We added the seal to the return pointers, so we had some way of figuring out whether two return pointers would actually fit together.}
\dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
\lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part. Specifically it is the part from the base to the current address of the stack pointer \emph{in $r_\stk$}. This means that if someone saves a stack pointer, so it is not included in the call, then they can still use it because that part of the stack was not encapsulated in a stack frame. Makes sense?}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

@@

\noindent\textbf{xjmp}\\
\lau{22-09-2017: Can we allow the callee to return any stack pointer? At the moment, it is not possible to create an empty stack pointer which may be a saving grace. If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
\begin{align*}
  \sem{\txjmp{r_1}{r_2}}(\Phi) = &
                                   \begin{cases}
                                     \arraycolsep=0pt
                                     \begin{array}[t]{rl}
                                       \Phi & \updReg{r_1}{w_1}\\
                                            & \updReg{r_2}{w_2}\\
                                            & \updReg{\pcreg}{c_\code}\\
                                            & \updReg{\rdata}{c_\data}
                                     \end{array} &
                                     \begin{array}[t]{l}
                                       \Phi(r_1) = \sealed{c_\code,\sigma} \tand \\
                                       \Phi(r_2) = \sealed{c_\data,\sigma} \tand \\
                                       \nonExec{c_\data} \tand \\
                                       w_1 = \linCons{\Phi(r_1)} \tand \\
                                       w_2 = \linCons{\Phi(r_2)}
                                     \end{array}
                                     \\&\\
                                     \sourcecolor\left.
                                       \arraycolsep=0pt
                                       \begin{array}[t]{rl}
                                         \Phi' & \updReg{r_2}{0} \\
                                               & \updReg{\pcreg}{\opc}\\
                                               & \updReg{\rdata}{0}\\
                                               & \updReg{\rstk}{c_\stk}
                                       \end{array}\right.& 
                                     \sourcecolor\left.
                                       \begin{array}[t]{l}
                                         \Phi(r_1) = \retptrc(\sigma) \tand \\
                                         \Phi(r_2) = \retptrd(\sigma) \tand \\
                                         \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\_} \tand \\
                                         \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
                                         \stkf = (\opc,\ms_{\stk,\priv},\aaddr_\stk) \tand \\
                                         \dom(\ms_{\stk,\priv}) = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
                                         \ms_{\stk,\priv}(\baddr_\stk-1) = \eaddr_\stk \tand \\
                                         c_\stk = \stkptr{\rw,\baddr_{\stk,\priv},\eaddr_\stk,\aaddr_\stk} \tand\\
                                         \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
                                       \end{array}\right.
                                     \\
                                     \failed & \totherwise
                                   \end{cases}
\end{align*}

@@

\noindent\textbf{call}\\
\lau{19-09-2017: How do we get the $\sigma_?$ for the return pointers? It should be in memory somewhere.}
\lau{19-09-2017: In which cases do we have undefined here?}
This instruction assumes that the current address of the stack pointer, $\aaddr_\stk$, points to the first free address of the stack. We assume a seal assignment function $\sealAss{}$ that given an address assigns a seal.\lau{25-09-2017: This will need some more explanation - namely, there should be some relationship between the available seals and this function (I also need to think more about the specifics of this).}
\begin{align*}
  \sem{\scall{\src{r_1}}{\src{r_2}}}(\Phi) = & 
                                               \begin{cases}
                                                 \Phi'
                                                 \arraycolsep=0pt
                                                 \begin{array}[t]{rl}
                                                   &\updReg{r_1}{w_1} \\
                                                        &\updReg{r_2}{w_2} \\
                                                        &\updReg{\pcreg}{c_1}\\
                                                        &\updReg{\rdata}{c_2}\\
                                                        &\updReg{\rstk}{c_\stk}\\
                                                        &\updReg{r_{\var{ret,c}}}{\retptrc(\sigma_?)}\\
                                                        &\updReg{r_{\var{ret,d}}}{\retptrd(\sigma_?)}
                                                 \end{array}
                                                 & 
                                                 \arraycolsep=0pt
                                                 \begin{array}[t]{l}
                                                   \Phi(r_1) = \sealed{c_1,\sigma_1} \tand \\
                                                   \Phi(r_2) = \sealed{c_2,\sigma_2} \tand \\
                                                   \sigma_1 = \sigma_2 \tand \\
                                                   \nonExec{c_2} \tand\\
                                                   \Phi = (\mem,\reg,\stk,\ms_\stk) \tand\\
                                                   \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} \tand \\
                                                   \baddr_\stk < \aaddr_\stk \leq \eaddr_\stk \tand \\
                                                   \ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\baddr,\aaddr_\stk - 1]} \tand \\
                                                   c_\stk = \stkptr{\baddr,\aaddr_\stk-1,\aaddr_\stk-1} \tand \\
                                                   \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                                   \opc = ((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen) \tand \\
                                                   \stk' = (\opc,\ms_{\stk,\priv},\aaddr_\stk) :: \stk \tand\\
                                                   \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
                                                   \mem(\baddr) = \seal{\_,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}\tand\\
                                                   \sigma_\baddr \leq \sigma_? \leq \sigma_\eaddr \tand\\
                                                   \sigma_? = \sealAss{a} \tand\\
                                                   w_1 = \linCons{\Phi(r_1)} \tand \\
                                                   w_2 = \linCons{\Phi(r_2)}
                                                 \end{array}
                                               \end{cases}
\end{align*}

>>>>

\newcommand{\scbnf}{\shareddom{sc}}
\newcommand{\cbnf}{\shareddom{c}}
\newcommand{\permbnf}{\shareddom{perm}}
\newcommand{\addrbnf}{\shareddom{a}}
\newcommand{\basebnf}{\shareddom{base}}
\newcommand{\aendbnf}{\shareddom{end}}
\newcommand{\rbnf}{\shareddom{r}}
%\newcommand{\glbnf}{\shareddom{gl}}
\newcommand{\linbnf}{\shareddom{l}}
\newcommand{\sealbasebnf}{\sigma_\shareddom{base}}
\newcommand{\sealendbnf}{\sigma_\shareddom{end}}

@@

%%% Constants
\newcommand{\constant}[1]{\mathrm{#1}}
\newcommand{\calllen}{\constant{call\_len}}
\newcommand{\stkb}{\constant{stk\_base}}

@@

% \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
% \lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
% \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf,\sigma}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma) \mid \retptrc(\sigma)}\\ 
    &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{06-09-2017: Should seals also have the option to be linear?}
\lau{13-09-2017: For now I have added linearity to the seals.}
\dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
\lau{28-09-2017: Why would you like this? If we want to add this, we will have to be careful, because an old code-return-pointer may fit perfectly well with our new data-return-pointer, so in this case it should not fail. We added the seal to the return pointers, so we had some way of figuring out whether two return pointers would actually fit together.}
\dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
\lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part. Specifically it is the part from the base to the current address of the stack pointer \emph{in $r_\stk$}. This means that if someone saves a stack pointer, so it is not included in the call, then they can still use it because that part of the stack was not encapsulated in a stack frame. Makes sense?}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\rbnf$ defines the finite set $\RegName$.
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

In both languages, the base address of the stack is known (the stack grows downwards in memory, so the base address marks the end of the stack). In the target language, this will be the base address of some linear capability. The base address can be any address on the machine, but it will have to remain the same. We write this constant as
\[
  \stkb
\]

@@

\noindent\textbf{xjmp}\\
\lau{22-09-2017: Can we allow the callee to return any stack pointer? At the moment, it is not possible to create an empty stack pointer which may be a saving grace. If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
\begin{align*}
  \sem{\txjmp{r_1}{r_2}}(\Phi) = &
                                   \begin{cases}
                                     \arraycolsep=0pt
                                     \begin{array}[t]{rl}
                                       \Phi & \updReg{r_1}{w_1}\\
                                            & \updReg{r_2}{w_2}\\
                                            & \updReg{\pcreg}{c_\code}\\
                                            & \updReg{\rdata}{c_\data}
                                     \end{array} &
                                     \begin{array}[t]{l}
                                       \Phi(r_1) = \sealed{c_\code,\sigma} \tand \\
                                       \Phi(r_2) = \sealed{c_\data,\sigma} \tand \\
                                       \nonExec{c_\data} \tand \\
                                       w_1 = \linCons{\Phi(r_1)} \tand \\
                                       w_2 = \linCons{\Phi(r_2)}
                                     \end{array}
                                     \\&\\
                                     \sourcecolor\left.
                                       \arraycolsep=0pt
                                       \begin{array}[t]{rl}
                                         \Phi' & \updReg{r_2}{0} \\
                                               & \updReg{\pcreg}{\opc}\\
                                               & \updReg{\rdata}{0}\\
                                               & \updReg{\rstk}{c_\stk}
                                       \end{array}\right.& 
                                     \sourcecolor\left.
                                       \begin{array}[t]{l}
                                         \Phi(r_1) = \retptrc(\sigma) \tand \\
                                         \Phi(r_2) = \retptrd(\sigma) \tand \\
                                         \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
                                         \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
                                         \stkf = (\opc,\ms_{\stk,\priv},\aaddr_\stk) \tand \\
                                         \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
                                         c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
                                         \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
                                       \end{array}\right.
                                     \\
                                     \failed & \totherwise
                                   \end{cases}
\end{align*}

@@

\noindent\textbf{call}\\
\lau{19-09-2017: How do we get the $\sigma_?$ for the return pointers? It should be in memory somewhere.}
\lau{19-09-2017: In which cases do we have undefined here?}
This instruction assumes that the current address of the stack pointer, $\aaddr_\stk$, points to the first free address of the stack. We assume a seal assignment function $\sealAss{}$ that given an address assigns a seal.\lau{25-09-2017: This will need some more explanation - namely, there should be some relationship between the available seals and this function (I also need to think more about the specifics of this).}
\begin{align*}
  \sem{\scall{\src{r_1}}{\src{r_2}}}(\Phi) = & 
                                               \begin{cases}
                                                 \Phi'
                                                 \arraycolsep=0pt
                                                 \begin{array}[t]{rl}
                                                   &\updReg{r_1}{w_1} \\
                                                        &\updReg{r_2}{w_2} \\
                                                        &\updReg{\pcreg}{c_1}\\
                                                        &\updReg{\rdata}{c_2}\\
                                                        &\updReg{\rstk}{c_\stk}\\
                                                        &\updReg{r_{\var{ret,c}}}{\retptrc(\sigma_?)}\\
                                                        &\updReg{r_{\var{ret,d}}}{\retptrd(\sigma_?)}
                                                 \end{array}
                                                 & 
                                                 \arraycolsep=0pt
                                                 \begin{array}[t]{l}
                                                   \Phi(r_1) = \sealed{c_1,\sigma_1} \tand \\
                                                   \Phi(r_2) = \sealed{c_2,\sigma_2} \tand \\
                                                   \sigma_1 = \sigma_2 \tand \\
                                                   \nonExec{c_2} \tand\\
                                                   \Phi = (\mem,\reg,\stk,\ms_\stk) \tand\\
                                                   \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} \tand \\
                                                   \baddr_\stk < \aaddr_\stk \leq \eaddr_\stk \tand \\
                                                   \ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\baddr_\stk,\aaddr_\stk - 1]} \tand \\
                                                   c_\stk = \stkptr{\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} \tand \\
                                                   \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                                   \opc = ((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen) \tand \\
                                                   \stk' = (\opc,\ms_{\stk,\priv},\aaddr_\stk) :: \stk \tand\\
                                                   \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
                                                   \mem(\baddr) = \seal{\_,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}\tand\\
                                                   \sigma_\baddr \leq \sigma_? \leq \sigma_\eaddr \tand\\
                                                   \sigma_? = \sealAss{a} \tand\\
                                                   w_1 = \linCons{\Phi(r_1)} \tand \\
                                                   w_2 = \linCons{\Phi(r_2)}
                                                 \end{array}
                                               \end{cases}
\end{align*}
